---
output: html_document
editor_options: 
  chunk_output_type: console
---
**Data Manipulation of greenhouse and climate data**
===

*Load libraries* IMPORTANT --> DO NOT SKIP
---
```{r, include=F}
library(dplyr)
library(plyr)
library(vegan)
library(usdm)
library(ggplot2)
library(ggfortify)
```

*Load data* IMPORTANT --> DO NOT SKIP
---
**NB: all traits are labeled 'T_' and all environmental variables are labeled 'E_' to easily distinguish between them in downstream analyses**
```{r}
#Phen.txt is Master_exp2a.xlsx
phenraw <-  read.table("phenraw.txt", header = T)

#SampleID	                Code for plant sample
#Mom	                    Code for maternal line
#Pop	                    Code for population of collection of mother
#Range	                  Continent of provenance
#ABCcluster               Cluster as defined for ABC analyses (UK = unknown, not used in ABC (native only, as introced ranges are grouped as a whole))
#Treatment	              Treatment of the sample. C = Control experiment a, watered 2x/day; D = Drought experiment a, watered on average every 1.5 day 
#T_fem_day	              First recorded day of female flowering (number of days after second transplant)
#T_male_day	              First recorded day of male flowering  (number of days after second transplant)
#T_pol_day	              First recorded day of pollen production (number of days after second transplant)
#T_seeds_day	            First recorded day of ripe seeds (number of days after second transplant)
#T_rac_die_day	          First recorded date of racemes dying  (number of days after second transplant)
#T_day_final    	        Date of final measurements (days after second transplant)
#T_die_day	              Number of days after second transplant no racemes produced pollen
#T_Branches	              Number of branches from main stem. Only branches ending in leaves are  counted
#T_Racemes	              number of racemes
#T_Raceme_length_longest	length of longest raceme. Did not take length of 'main' raceme, as in many cases discrimination was impossible
#T_Longest_leaf	          Length of longest leaf in mm. Longest leaf might have died/fallen off due to scenescence
#T_width_longest_leaf	    width of longest leaf in mm. Not always possible due to scenescence. 
#T_Stem_width_mum	        width of stem at base, in micro meter
#T_Height	                height of plant in mm
#T_Width	                width of plant in mm
#T_angle	                Average branch angle to main stem. Moving a protractor along the main stem and reading off branch angles. An average   measure of angle if #number of branches was more than 2 and variation of mean angle did not exceed 10 degrees 
#T_male_weight	          Dry weight of racemes in milligrams
#T_plant_weight	          Dry weight of above-ground biomass in grams, including seeds, excluding racemes
#T_plant_noseed	          Dry weight of above-ground biomass in grams, excluding seeds, excluding racemes
#T_root_weights	          Dry weight of below-ground biomass in grams
#T_seed_weight_tot	      Dry total weight of all seeds in milligrams
#T_seed_20	              Dry weight of 20 seeds in milligrams
#T_leaf_weight	          Dry weight of fully-expanded new leaf in milligrams
#T_leaf_nostem	          Dry weight of above leaves without the stalk
#T_leaf_area	            Area of leaf in square cm *10
##T_leaf_perim	          Perimeter of leaf in cm *10
#T_leaf_image           	Type of image used to save leaf area photos (calcutions from png images seemed dubious)
#T_germ	                  Proportion of maternal seeds germinated
#T_momseed_mg	            Average seed weight of maternal line

# reformat data type
phenraw$Treatment <- as.factor(phenraw$Treatment) 
phenraw$T_seed_weight_tot <- as.numeric(phenraw$T_seed_weight_tot) 
phenraw$T_seed_20 <- as.numeric(phenraw$T_seed_20)

# Further calculations per trait category

#Size
phenraw$T_shoot_biom <- phenraw$T_male_weight + (phenraw$T_plant_weight)*1000 # the latter also includes seeds
phenraw$T_tot_biom <- (phenraw$T_shoot_biom + phenraw$T_root_weights*1000)/1000

#phenology
#phenraw$T_fem_length <- phenraw$T_seeds_day - phenraw$T_fem_day # length of female flowering (day of ripe seeds - day of first occurance)
#phenraw$T_male_length <- phenraw$T_rac_die_day - phenraw$T_pol_day # length of male flowering (day of racemes dying - day of first pollen production)
phenraw$T_fl_end<-apply(phenraw[,c("T_seeds_day","T_rac_die_day")],1,max) # last day of either first occurance ripe seeds or first day racemes daying
phenraw$T_fl_start <- apply(phenraw[,c("T_fem_day","T_pol_day")],1,min) # first day of either first occurance female or first pollen production
#phenraw$T_flower_length <- phenraw$T_fl_end - phenraw$T_fl_start # length of flowering season
phenraw$T_sexmismatch <- phenraw$T_pol_day - phenraw$T_fem_day
#phenraw$T_sexmismatch_length <- phenraw$T_male_length - phenraw$T_fem_length

# Fitness
phenraw$T_seeds_tot <- (phenraw$T_seed_weight_tot/ phenraw$T_seed_20) * 20 # total number of seeds per plant --> this is a total fitness measure
phenraw$T_fit_all <- (phenraw$T_male_weight + phenraw$T_seed_weight_tot)/1000
phenraw$T_repal <- phenraw$T_fit_all / phenraw$T_tot_biom
#henraw$T_repalm <- phenraw$T_male_weight / phenraw$T_tot_biom
#phenraw$T_repalf <- phenraw$T_seed_weight_tot / phenraw$T_tot_biom

# Physiological traits
phenraw$T_leaf_weight <- as.numeric(phenraw$T_leaf_weight) 
phenraw$T_Sla <- (phenraw$T_leaf_area*100) / phenraw$T_leaf_weight # Specific Leaf Area ((mm^2)/mg)

#Relative investment
phenraw$T_sexrat_weight <- phenraw$T_seed_weight_tot /phenraw$T_male_weight # sex-ratio expressed in biomass of reproductive structures
#phenraw$T_sexrat_numweight <- phenraw$T_seeds_tot / phenraw$T_male_weight # sex-ratio expressed in number of female / biomass of male reproductive structure
phenraw$T_rootshoot_rat <- phenraw$T_root_weights*1000 / phenraw$T_shoot_biom # root/shoot ratio

#remove traits which are inconsistent or dubious
phenraw <- phenraw[ ,-which(names(phenraw) %in% c("T_Width" # highly inconsisten according to growth data
                                         ,"T_angle" # inconsistent measurements
                                         ,"T_leaf_image" # removed data retrieved from .png images before loading datafile
                                         ,"T_Height" # Maximum height occasionaly occured before final measurement, get from growth file below
                                         ,"T_leaf_nostem" # only measured for some leaves, highly linear with leaf_weight, so doesn't add info
                                         ,"T_plant_weight" #incorporated in shoot_biom
                                         ,"T_plant_noseed" # incorporated in shoot_biom
                                         ,"T_width_longest_leaf" #highly colinear with length longest leaf
                                         ))]

#remove traits which are unnecessary or have a better descriptive trait; physiology
phenraw <- phenraw[ ,-which(names(phenraw) %in% c("T_leaf_weight" # better measure is SLA; variable dependent on plant size
                                         ,"T_leaf_area" # better measure is SLA; variable dependent on plant size
                                         ,"T_leaf_perim" # variable dependent on leaf/plant size. Inconsistency in data collection, can be improved
                                         ))]

#remove traits which are unnecessary or have a better descriptive trait; phenrawology
phenraw <- phenraw[ ,-which(names(phenraw) %in% c("T_seeds_day" # already captured by fem_length
                                         ,"T_rac_die_day" # already captured by mal_length
                                         ,"T_fl_end" # already captured by fl_length & flower_start
                                         ,"T_day_final" # already captured by fl_length & flower_start
                                         ,"T_stop_day" # already captured by fl_length & flower_start
                                         ,"T_male_day" #already captured by pol
                                         ))]

#remove traits which are unnecessary or have a better descriptive trait; size
phenraw <- phenraw[ ,-which(names(phenraw) %in% c("T_minHeight_cm" # determined by germ
                                         ,"T_root_weights" # captured by root/shoot ratio; trait dependend on plant size
                                                                                 ))]
# remove sample with high number of inconsistencies
phenraw <- phenraw[-which(phenraw$SampleID=="AU23-8F"),]
```

```{r}
# remove duplicate samples within treatments with same maternal line (manually to ensure same samples are used as for sequencing)
phenraw = phenraw[!phenraw$SampleID %in% c("AA2B-5C","AU1-6D","AU1-8C","AU4-15D","AL-25C","AU5-19A","EU1-16C","EU11-30B","EU32-13D","AU11-1B","AU11-26B","AU11-5A","AU12-6B","AU13-1D","AU13-17A","EU14-10B","EU14-5A","EU26-26C","EU30-1A","AU15-21A","AU18-14B","AU23-1B","AU27-16A","AU27-2A","AU33-9C","EU15-2A","EU15-4D","EU15-8A","EU22-8D","GA-24C","MA1-10C","GA-27B","GA-27C","GA-27D","PA-17A","PA-17D"),]


# remove populations with <4 samples within treatments
phenraw = phenraw[!phenraw$Pop %in% c("AA20","AU11"),]
#phenraw <- phenraw[-which(phenraw$Pop=="AA20"),]

# remove populations which have not enough individuals in drought treatment only
phenraw <- phenraw %>%
  filter(!(Treatment=="D" & Pop == "AU09")) %>%
  filter(!(Treatment=="D" & Pop == "EU05")) %>%
  filter(!(Treatment=="D" & Pop == "EU07")) %>%
  filter(!(Treatment=="D" & Pop == "FL")) 

# remove samples not replicated in each treatment per maternal line (remove all Mom's with count <2)
phen =  phenraw[phenraw$Mom %in% names(table(phenraw$Mom))[table(phenraw$Mom) > 1],]
```

Read in other datasets
```{r}
growth <- read.csv("IndivModOut_161210.csv", header = T) # from Growth script
growth <-growth[,-c(2,4:20)] # remove all non-growth variables (t- and p-val's etc)

#   - Maximum growth rate                       (MaxSlope_cm_Day)
#   - Budbreak day                              (Growth5pct_Day)
#   - Height at 5% growth i.e. Budbreak,        (Growth5pct_cm)
#   - Budset day                                (Growth95pct_Day)
#   - Height at 95% growth i.e. Budset,         (Growth95pct_cm)
#   - Length of growing period                  (Growth5-95pct_Days)
#   - Height gain during growing period         (Growth5-95pct_cm)
#   - Day that maximum growth rate starts       (LinGrStart_Day)
#   - Day that maximum growth rate ceases       (LinGrEnd_Day)
#   - Duration of maximum growth rate           (LinearGrowth_Days)
#   - Day at which half annual growth occurs    (InflxPt_Day)
#   - Height attained at half annual growth     (InflxPt_Ht_cm)

phen <- left_join(phen, growth, by="SampleID")
phen$T_MaxSlope_cm_Day <- phen$T_MaxSlope_cm_Day/10
phen$T_maxHeight_cm <- phen$T_maxHeight_cm/10

struc<- read.table("struc_ddgbsall.txt", header=T)
phen<- left_join(phen, struc, by = "Pop")
noli <- read.table("phenol.txt",header=T)
phen <- left_join(phen, noli, by="SampleID")
Ho <- read.table("Ho.txt", header=T)
```

```{r}
climfile <- read.table("clim.txt", header=T)
 # bioclim 1-19 & lat,lon, alt for all sampling locations, not standardised (preliminary analysis showed standardisation was not necessary)
#names(climfile) <- gsub(x = names(climfile), pattern = "\\E_", replacement = "")  
#names(climfile) <- gsub(x = names(climfile), pattern = "\\_all", replacement = "")  

# BIO1 = Annual Mean Temperature
# BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
# BIO3 = Isothermality (BIO2/BIO7) (* 100)
# BIO4 = Temperature Seasonality (standard deviation *100)
# BIO5 = Max Temperature of Warmest Month
# BIO6 = Min Temperature of Coldest Month
# BIO7 = Temperature Annual Range (BIO5-BIO6)
# BIO8 = Mean Temperature of Wettest Quarter
# BIO9 = Mean Temperature of Driest Quarter
# BIO10 = Mean Temperature of Warmest Quarter
# BIO11 = Mean Temperature of Coldest Quarter
# BIO12 = Annual Precipitation
# BIO13 = Precipitation of Wettest Month
# BIO14 = Precipitation of Driest Month
# BIO15 = Precipitation Seasonality (Coefficient of Variation)
# BIO16 = Precipitation of Wettest Quarter
# BIO17 = Precipitation of Driest Quarter
# BIO18 = Precipitation of Warmest Quarter
# BIO19 = Precipitation of Coldest Quarter

climfile <- climfile[,-which(names(climfile)%in% c(#"E_Longitude", # not descriptive of anything
                                                   "E_dem_all" # correlated with alt_all, but from different database
                                                   ,"Range" #will be added later
                                                   ))]

# Remove populations not in the dataset
clim <- semi_join(climfile, phen, by="Pop")
```


*Remove variables from dataframe as per above methods* IMPORTANT --> DO NOT SKIP
---
```{r}
phen <- phen[ ,-which(names(phen) %in% c("T_InflxPt_Ht_cm", "T_LinGrStart_Day", "T_LinGrEnd_Day",  "T_Growth5pct_cm", "T_Growth5pct_Day", "T_Growth95pct_cm", "T_Growth95pct_Day", "T_Growth5.95pct_cm", "T_Growth5.95pct_Days"))]
phen_ho <- inner_join(phen, Ho, by= "SampleID")
```

*Aggregate by Treatment and Population ID to get averages* IMPORTANT --> DO NOT SKIP
---
```{r}
#taking out too much information resulting in downstream issues with mvpart
agg_phen_mean <- aggregate(phen, by= list(phen$Treatment, phen$Pop), FUN=mean, na.rm=T)
colnames(agg_phen_mean)[1:2] <- c("Treatment", "Pop")
d <- phen[,3:4] #Get Pop and Range columns
d <- distinct(d) # remove all duplicate rows
agg_phen_mean <- agg_phen_mean[,-c(3:7)] # remove empty columns
phen_mean <- left_join(agg_phen_mean,d, by="Pop") # add Range info to dataframe joining by pop

agg_phenho_mean <- aggregate(phen_ho, by= list( phen_ho$Pop), FUN=mean, na.rm=T)
colnames(agg_phenho_mean)[1] <- c("Pop")
d <- phen_ho[,3:4] #Get Pop and Range columns
d <- distinct(d) # remove all duplicate rows
agg_phenho_mean <- agg_phenho_mean[,-c(2:6)] # remove empty columns
phenho_mean <- left_join(agg_phenho_mean,d, by="Pop") # add Range info to dataframe joining by pop
```

*Join the phenotypic and climatic data* IMPORTANT --> DO NOT SKIP
---
```{r}
pc_mean <- left_join(phen_mean, clim, by="Pop")
pcsel_mean <- left_join(phen_mean, climsel, by="Pop")
phencl <- left_join(phen, clim, by="Pop")
pcho_mean <- left_join(phenho_mean, clim, by="Pop")
```

*Calculate PC scores for environmental variables*
---
See exploreGH for elaborate PCA approach
```{r}
#with standarisation --> better solution for climate, without latitude
climnolatnolonst_pca <- prcomp((pc_mean[,54:74]+500), scale=T)
climlatst_pca <- prcomp((pc_mean[,c(52:74)]+500), scale=T)

autoplot(climnolatnolonst_pca,	type="obs")

pdf("_Graph/_Greenhouse/PCA_ranges_nolatlon.pdf", height=5, width=8)
autoplot(climnolatnolonst_pca, data=pc_mean, colour="Range",label.label="Pop",loadings=T,label.size=3, loadings.label=T,loadings.colour=1,loadings.label.colour=1,loadings.label.angle=-10,loadings.label.hjust=-.1,shape=F,loadings.label.repel=T,  loadings.label.size=3,frame = TRUE, frame.alpha=.05)+theme_bw()
dev.off()

pdf("_Graph/_Greenhouse/PCA_ranges.pdf", height=5, width=8)
autoplot(climlatst_pca, data=pc_mean, colour="Range",label.label="Pop",loadings=T,label.size=3, loadings.label=T,loadings.colour=1,loadings.label.colour=1,loadings.label.angle=-10,loadings.label.hjust=-.1,shape=F,  loadings.label.size=3,frame = TRUE, frame.alpha=.05)+theme_bw()
dev.off()

aload <- abs(climlatst_pca$rotation)
sweep(aload, 2, colSums(aload), "/")
summary(climnolatnolonst_pca)

pc_mean$PC1 <- climnolatst_pca$x[,1]
pc_mean$PC2 <- climnolatst_pca$x[,2]
pc_mean$PC3 <- climnolatst_pca$x[,3]
```

*Write datasets*
---
```{r}
write.table(phen,"phen.txt")
write.table(phen_ho,"phen_ho.txt")
write.table(pc_mean,"pc_mean.txt")
write.table(pcho_mean,"pcho_mean.txt")
write.table(pcsel_mean,"pcsel_mean.txt")
write.table(phencl,"phencl.txt")
```


**END SCRIPT**