---
output: html_document
editor_options: 
  chunk_output_type: console
---

**Step 0) Data Manipulation**
*a) Load libraries* IMPORTANT --> DO NOT SKIP
---
```{r, include=F}
library(dplyr)
library(plyr)
library(vegan)
library(usdm)
library(ggfortify)
library(ggplot2)
library(nlme)
library(lmerTest)
library(lme4)
library(mvpart) # for MRT, package needs to be installed from github using devtools::install_github("cran/mvpart")
library(visreg)
library(gridExtra)
library(multcomp)
library(phia)
library(Hmisc) # for rcorr

```


*Load data* IMPORTANT --> DO NOT SKIP
---
```{r}
phen <- read.table("phen.txt", header=T)
pc_mean <-read.table("pc_mean.txt", header=T)
clim = read.table("clim.txt", header=T)
```

*Create subsets* IMPORTANT --> DO NOT SKIP
---
```{r}
# individual trait data per Treatment
phenC <- phen[which(phen$Treatment == 'C'),]
phenD <- phen[which(phen$Treatment == 'D'),]

phen_mean_C <- phen_mean[which(phen_mean$Treatment=="C"),]
phen_mean_D <- phen_mean[which(phen_mean$Treatment=="D"),]

# individual trait data per Treatment per continent
phenC_NA <- phenC[which(phenC$Range == 'Native'),]
phenC_EU <- phenC[which(phenC$Range == 'Europe'),]
phenC_AU <- phenC[which(phenC$Range == 'Australia'),]
phenD_NA <- phenC[which(phenD$Range == 'Native'),]
phenD_EU <- phenC[which(phenD$Range == 'Europe'),]
phenD_AU <- phenC[which(phenD$Range == 'Australia'),]

pc_mean_C <- pc_mean[which(pc_mean$Treatment=="C"),]
pc_mean_D <- pc_mean[which(pc_mean$Treatment=="D"),]
```
  
**2) Trait correlation matrices** CHECK COLUMNS INCLUDED BEFORE RUNNING
*a) Calculate correlations between all variables*
---

Function to transform rcorr output
```{r}
# function to flatten rcorr matrix
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```

Using Bonferroni correction to account for multiple comparisons (multiplies p-val times number of comparisons made). Use data without transformed variables as Spearman's correlation test doesn't need it
```{r}
#All, Control treatment
pc_mean_sub <- pc_mean[3:36]
corrs <- rcorr(as.matrix(pc_mean_sub), type="spearman")

corrs_r <- data.frame(corrs$r)
corrs_p <- data.frame(corrs$P)
corrs_pbonf <- data.frame(corrs$p_bonf)
corr <- flattenCorrMatrix(corrs_r,corrs_p)
corr$p_bonf <- p.adjust(corr$p, method= "bonferroni")
write.table(corr, "correlations_all_phenmean.txt")  
```  
  
```{r}
clim_sub = pc_mean[48:69]
corrs <- rcorr(as.matrix(clim_sub), type="spearman")

corrs_r <- data.frame(corrs$r)
corrs_p <- data.frame(corrs$P)
corrs_pbonf <- data.frame(corrs$p_bonf)
corr <- flattenCorrMatrix(corrs_r,corrs_p)
corr$p_bonf <- p.adjust(corr$p, method= "bonferroni")
write.table(corr, "correlations_all_clim.txt")  
```

*Calculate PCA for clim*
---
```{r}
colnames(clim) <- gsub("_all","",colnames(clim))
colnames(clim) <- gsub("E_","",colnames(clim))
colnames(clim) <- gsub("Latitudeabs","Lat",colnames(clim))
colnames(clim) <- gsub("Longitude","Long",colnames(clim))
colnames(clim) <- gsub("alt","Alt",colnames(clim))

colnames(pc_mean_C) <- gsub("_all","",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("E_","",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("Latitudeabs","Lat",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("Longitude","Long",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("alt","Alt",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio10","TQwarm",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio11","TQcold",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio12","Pmean",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio13","Pmax",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio14","Pmin",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio15","Pseas",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio16","PmaxQ",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio17","PminQ",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio18","PQwarm",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio19","PQcold",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio1","Tmean",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio2","Tdiur",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio3","Tiso",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio4","Tseas",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio5","Tmax",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio6","Tmin",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio7","Trange",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio8","TQwet",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("bio9","TQdry",colnames(pc_mean_C))

```

```{r}
clim[,c(3:4,6:25)] = apply(clim[,c(3:4,6:25)], 2, function(x) as.numeric(as.character(x)));

clim_pca<- prcomp(clim[c(3:4,6:25)], scale=T)

autoplot(clim_pca, data=clim[c(2:4,6:25)], colour="Range", loadings=T, loadings.label=T, shape=F, label.label="Pop", loadings.label.size=4)+ theme_minimal()  

pdf("_Graph/PCA_clim.pdf", height=6, width=8)
autoplot(clim_pca, data=clim[c(2:4,6:25)], colour="Range", loadings=T, loadings.label=T, shape="Range", loadings.label.size=4, loadings.label.repel=T, loadings.label.colour='black')+ theme_minimal() + scale_shape_manual(values=c(19,15,17))
dev.off()

aload <- abs(clim_pca$rotation)
sweep(aload, 2, colSums(aload), "/")

climsel_pca <- prcomp(wisconsin(pc_mean_C[c(48:49,55,57,62,64:65)]+500), scale=T)
autoplot(climsel_pca, data=pc_mean_C, colour="Range", loadings=T, loadings.label=T, shape=F, label.label="Pop", loadings.label.size=4)+ theme_minimal()  

pdf("_Graph/PCA_topPEA.pdf", height=6, width=8)
autoplot(climsel_pca, data=pc_mean_C, colour="Range", loadings=T, loadings.label=T, shape="Range", loadings.label.size=4, loadings.label.repel=T, loadings.label.colour='black')+ theme_minimal() + scale_shape_manual(values=c(19,15,17))
dev.off()
aload <- abs(climsel_pca$rotation)
sweep(aload, 2, colSums(aload), "/")
summary(climsel_pca)
```

*d)	Calculate final PC for traits**
---
```{r}
colnames(pc_mean_C) <- gsub("T_tot_biom","Biom",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_Branches","Branch",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_maxHeight_cm","Height",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_rootshoot_rat","RS",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_Sla","SLA",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_Stem_width_mum","Stem",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_sexmismatch","Dich",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_sexrat_weight","Sex",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_fl_start","Flow",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_MaxSlope_cm_Day","Grow",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_InflxPt_Day","Half",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_fit_all","Fit",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_male_weight","Male",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_repal","Rfit",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_seed_weight_tot","Seed",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("T_seed_20","Ssize",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("PeakArea","Area",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("Peaks","Peaks",colnames(pc_mean_C))
colnames(pc_mean_C) <- gsub("Trichome","Trich",colnames(pc_mean_C))
```


```{r}
phen_pca <- prcomp(na.omit(pc_mean_C[c(13,17:19,21:24,26:27)]+100), scale=T) 

aload <- abs(phen_pca$rotation)
sweep(aload, 2, colSums(aload), "/")

pdf("_Graph/PCA_traits.pdf", height=6, width=8)
autoplot(phen_pca, na.omit(pc_mean_C[c(6,1:13,17:19,21:28,35,45:47)]), colour="Range", loadings=T, loadings.label=T, shape="Range", loadings.label.size=4, loadings.label.repel=T, loadings.label.colour='black')+ theme_minimal() + scale_shape_manual(values=c(19,15,17))
dev.off()

phen_pcasub <- prcomp(na.omit(pc_mean_C[c(6,11,17:19,24,26,28,47)]+100), scale=T) 

pdf("_Graph/PCA_topPEAtraits.pdf", height=6, width=8)
autoplot(phen_pcasub, na.omit(pc_mean_C[c(6,11,17:19,24,26,28,47)]), colour="Range", loadings=T, loadings.label=T, shape="Range", loadings.label.size=4, loadings.label.repel=T, loadings.label.colour='black')+ theme_minimal() + scale_shape_manual(values=c(19,15,17))
dev.off()
aload <- abs(phen_pcasub$rotation)
sweep(aload, 2, colSums(aload), "/")
```

**END SCRIPT**
===