
*Load libraries* IMPORTANT --> DO NOT SKIP
---
```{r, include=F}
library(knitr)
library(dplyr)
library(plyr)
library(vegan)
library(usdm)
library(ggfortify)
library(nlme)
library(lmerTest)
library(lmtest)
library(lme4)
library(mvpart) # for MRT, package needs to be installed from github using devtools::install_github("cran/mvpart")
library(visreg)
library(gridExtra)
library(multcomp) #glht
library(phia) #testInteractions
library(ape) #varcomp
library(Hmisc) # for rcorr
library(reshape)
opts_knit$set(root.dir = '/Users/lottevanboheemen/Documents/Monash/PhD/Analyses/Data', warning=FALSE, message=T)
```

*Load data* IMPORTANT --> DO NOT SKIP
---
```{r}
phen <- read.table("phen.txt", header=T) #non-transformed individual traits
pc_mean <-read.table("pc_mean.txt", header=T) #population mean traits + bioclimatic variables
pcho_mean<- read.table("pcho_mean.txt",header=T)
```

*Create subsets* IMPORTANT --> DO NOT SKIP
---

```{r}
phenC <- phen[which(phen$Treatment == 'C'),]
phenD <- phen[which(phen$Treatment == 'D'),]

pc_meanC <- pc_mean[which(pc_mean$Treatment == 'C'),]
pc_meanD <- pc_mean[which(pc_mean$Treatment == 'D'),]
```

*Test the trait differences between Ranges in MANOVA*
---

Do this without any rho>0.7 traits
```{r}
m.man<-manova(cbind( T_fl_start, T_repal, sqrt(T_Sla), T_fit_all,sqrt(T_sexmismatch+60), sqrt(T_seed_20),  T_MaxSlope_cm_Day) ~ Range*E_Latitudeabs+I(E_Latitudeabs^2), data=pc_meanC)
summary(m.man, test="Wilks")

summary.aov(m.man, test="Wilks") #every trait effect

testInteractions(m.man, pairwise = "Range", test="Wilks")
testInteractions(m.man, pairwise = "Range",slope="E_Latitudeabs", test="Wilks")
```


*Univariate phenotypic responses to range and latitude*
---

```{r}
#maxHeight
step(lmer(sqrt(T_maxHeight_cm)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m<-(lmer(sqrt(T_maxHeight_cm) ~  E_Latitudeabs *Range + (1 | q), data = pc_meanC))
visreg(m, by="Range", xvar="E_Latitudeabs", overlay=T)

interactionMeans(m, factor="Range", slope = "E_Latitudeabs")
testInteractions(m, pairwise = "Range", slope = "E_Latitudeabs")

#test if most northern Australian plants differ significantly from the native range in size
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Lowest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Lowest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Highest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Lowest EU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Highest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Highest EU lat


#Total biomass
step(lmer(sqrt(T_tot_biom)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m<- lmer(sqrt(T_tot_biom) ~  Range*E_Latitudeabs +I(E_Latitudeabs^2)+ (1 | q) , data = pc_meanC)
interactionMeans(m, factor="Range", slope= "E_Latitudeabs")
testInteractions(m, pairwise = "Range", slope = "E_Latitudeabs")

testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Lowest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Lowest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Highest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Lowest EU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Highest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Highest EU lat


#Max. growth rate
step(lmer(T_MaxSlope_cm_Day~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), keep.effs="E_Latitudeabs",reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m = lmer((T_MaxSlope_cm_Day) ~Range+E_Latitudeabs+ (1|q), data = pc_meanC)
testInteractions(m, "Range")


#Flowering onset
step(lmer(T_fl_start~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
Anova(lmer(T_fl_start~ Range+E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger", test.statistic = "F")
Anova(lmer(T_fl_start~ Range+E_Latitudeabs+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger", test.statistic = "F")

m <- lmer((T_fl_start)~ Range+E_Latitudeabs+ I(E_Latitudeabs^2) + (1 | q), data = pc_meanC)
testInteractions(m,"Range")

#Dichogamy
step(lmer(sqrt(T_sexmismatch+60)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
step(lmer(sqrt(T_sexmismatch+60)~ Range+E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m <- lmer(sqrt(T_sexmismatch+60) ~  E_Latitudeabs+Range +I(E_Latitudeabs^2)+ (1 | q), data = pc_meanC)
testInteractions(m,"Range")


#Floral sex allocation (female/male)
step(lmer(log(T_sexrat_weight)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m <- lmer(log(T_sexrat_weight) ~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1 | q)  , data = pc_meanC)
s= interactionMeans(m)
r= exp(s$`adjusted mean`)
se=r= exp(s$`adjusted mean`)

(r[1]-r[3])
(r[2]-r[3])

testInteractions(m,pairwise = "Range", slope = "E_Latitudeabs")

testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Lowest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Lowest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Highest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Lowest EU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Highest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Highest EU lat

#Weight per seed
step(lmer(sqrt(T_seed_20)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m <- lmer(sqrt(T_seed_20) ~ Range+ E_Latitudeabs +   I(E_Latitudeabs^2) + (1 | q),  data = pc_meanC)

testInteractions(m,"Range")

#Total reproductive biomass
step(lmer((T_fit_all)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
step(lmer((T_fit_all)~ Range+E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m <- lmer((T_fit_all)~ I(E_Latitudeabs^2)+Range+E_Latitudeabs+ (1|q), data = pc_meanC)
testInteractions(m,"Range")


#Relative reproductive biomass
step(lmer((T_repal)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m <- lmer((T_repal) ~ Range *E_Latitudeabs + (1 | q) ,data = pc_meanC)
visreg(m, by="Range",xvar="E_Latitudeabs")

testInteractions(m,pairwise = "Range", slope = "E_Latitudeabs")
interactionMeans(m, factor="Range", slope = "E_Latitudeabs")

testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Lowest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Lowest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Australia")]))) #Highest AU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=min(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Lowest EU lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Native")]))) #Highest NA lat
testInteractions(m, pairwise = "Range", covariates = c(E_Latitudeabs=max(pc_mean$E_Latitudeabs[which(pc_mean$Range=="Europe")]))) #Highest EU lat


#SLA
step(lmer(sqrt(T_Sla)~ Range*E_Latitudeabs+I(E_Latitudeabs^2)+ (1|q) , data = pc_meanC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

m<- lmer(sqrt(T_Sla)~ Range+E_Latitudeabs+ (1 | q), data = pc_meanC)
testInteractions(m, "Range")
```


*Height-flowering patterns*
---

Indivq as I testing the same populations

```{r}
##Plant height in relation to phenology
step(lmer(T_fl_starttr~(T_maxHeight_cm_sq)*Range + (1|Pop) + (1|q), data=phentrC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
m <- lmer(T_fl_start~sqrt(T_maxHeight_cm)*Range + (1|Pop), data=phenC)
visreg(m, xvar="T_maxHeight_cm", by= "Range" , partial=T, overlay=T, ylab="Start of flowering (days after T2)", xlab="Maximum height (mm)",cex.lab=1.5)
interactionMeans(m, factor = "Range", slope="sqrt(T_maxHeight_cm)")
testInteractions(m,slope="sqrt(T_maxHeight_cm)")

##Plant height in relation to dichogamy
phentrC$sm60lg <- log(phentrC$T_sexmismatch+60) #hack to be able to compare adjusted slopes. 
step(lmer(sm60lg~(T_maxHeight_cm_sq)*Range+ (1|Pop) +(1|q), data=phentrC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
m <- lmer(log(T_sexmismatch+60)  ~ sqrt(T_maxHeight_cm)*Range + (1 | Pop),  data = phenC)
visreg(m, xvar="T_maxHeight_cm", by= "Range",trans=function(x) exp(x)-60,partial=T, overlay=T, ylab="Dichogamy \n(days, female-male flowering)", xlab="Maximum height (mm)",cex.lab=1.5)
interactionMeans(m,factor = "Range", slope="sqrt(T_maxHeight_cm)") #slopes in each range
testInteractions(m, pairwise = "Range", slope="sqrt(T_maxHeight_cm)")

##Plant height in relation to sex allocation
step(lmer(T_sexrat_weight_lg~(T_maxHeight_cm_sq)*Range + (1|Pop) + (1|q), data=phentrC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
m <- lmer(log(T_sexrat_weight+1)~ sqrt(T_maxHeight_cm)*Range  + (1 | Pop), data = phenC)
visreg(m, xvar="T_maxHeight_cm", by= "Range",partial=T, overlay=T, ylab="Sex allocation log(female/male biomass+1)", xlab="Maximum height (mm)",cex.lab=1.5)
interactionMeans(m,factor = "Range", slope="sqrt(T_maxHeight_cm)") 
testInteractions(m, pairwise = "Range", slope="sqrt(T_maxHeight_cm)")

##Plant height in relation to seed weight
step(lmer(T_seed_20_lg~(T_maxHeight_cm_sq)*Range + (1|Pop), data=phentrC), reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
m <- lmer(log(T_seed_20)~sqrt(T_maxHeight_cm)*Range + (1 | Pop), data = phenC)
visreg(m, xvar="T_maxHeight_cm", by= "Range" , trans= function(x) x^2, partial=T, overlay=T, ylab="Seed weight", xlab="Maximum height (mm)")
interactionMeans(m, factor = "Range", slope="sqrt(T_maxHeight_cm)") 
testInteractions(m,pairwise="Range", slope="sqrt(T_maxHeight_cm)")# 
```

*Heterozygosity*
---

Heterozygosity-latitude relationship
```{r}
plot(pcho_mean$Ho~pcho_mean$E_Latitudeabs)
Anova(lmer(Ho~Range*E_Latitudeabs+(1|q), data=pcho_mean),type=3)
Anova(lm(Ho~Range*E_Latitudeabs, data=pcho_mean),type=3)
Anova(lmer(Ho~Range*E_Latitudeabs+Range*I(E_Latitudeabs^2) + (1|q), data=pcho_mean),type=3)
Anova(lmer(Ho~Range*E_Latitudeabs+I(E_Latitudeabs^2) + (1|q), data=pcho_mean),type=3)
Anova(lmer(Ho~Range+E_Latitudeabs+I(E_Latitudeabs^2) + (1|q), data=pcho_mean),type=3)
Anova(lmer(Ho~Range+E_Latitudeabs + (1|q), data=pcho_mean),type=3)
Anova(lmer(Ho~Range + (1|q), data=pcho_mean),type=3)
m <- lmer(Ho~Range + (1|q), data=pcho_mean)
testInteractions(m)
```


Inluding latitude effect, testing on population means
```{r}
Anova(lmer(sqrt(T_shoot_biom)~ Ho*Range+E_Latitudeabs*Range +I(E_Latitudeabs^2)+ (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")
Anova(lmer(sqrt(T_shoot_biom)~ Ho*Range+E_Latitudeabs +I(E_Latitudeabs^2)+ (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")
Anova(lmer(sqrt(T_shoot_biom)~ Ho*Range+E_Latitudeabs + (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")

m<-lmer(sqrt(T_tot_biom/1000)~ Ho*Range+E_Latitudeabs + (1|q) , data = pcho_mean)
visreg(m, by="Range",xvar="Ho", overlay=T)
plot(interactionMeans(m, factor="Range", slope="Ho"))
testInteractions(m, pairwise = "Range", slope="Ho")
m1<-lm(sqrt(T_tot_biom/1000)~ Ho*Range+E_Latitudeabs , data = pcho_mean)
visreg(m1, by="Range",xvar="Ho", overlay=T,xlab="Observed heterozygosity", ylab="Total biomass (g)",  trans=function(x) x^2, partial=T)
interactionMeans(m1, factor="Range", slope="Ho")
testInteractions(m, pairwise = "Range", slope="Ho")

plot(effect("Ho", m1, partial.residuals = TRUE))


Anova(lmer(sqrt(T_tot_biom)~ Ho*Range+E_Latitudeabs*Range +I(E_Latitudeabs^2)+ (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")
Anova(lmer(sqrt(T_tot_biom)~ Ho*Range+E_Latitudeabs +I(E_Latitudeabs^2)+ (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")
Anova(lmer(sqrt(T_tot_biom)~ Ho*Range+E_Latitudeabs + (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")

m<-lmer(sqrt(T_tot_biom/1000)~ Ho*Range+E_Latitudeabs + (1|q) , data = pcho_mean)
visreg(m, by="Range",xvar="Ho", overlay=T)
plot(interactionMeans(m, factor="Range", slope="Ho"))
testInteractions(m, pairwise = "Range", slope="Ho")
m1<-lm(sqrt(T_tot_biom/1000)~ Ho*Range+E_Latitudeabs , data = pcho_mean)
visreg(m1, by="Range",xvar="Ho", overlay=T,xlab="Observed heterozygosity", ylab="Total biomass (g)",  trans=function(x) x^2, partial=T)
interactionMeans(m1, factor="Range", slope="Ho")
testInteractions(m, pairwise = "Range", slope="Ho")

plot(effect("Ho", m1, partial.residuals = TRUE))

step(lmer(sqrt(T_fit_all) ~ Ho*Range+E_Latitudeabs*Range +I(E_Latitudeabs^2)+ (1|q), data= pcho_mean),  keep.effs="Ho", reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")


step(lmer(log(T_seed_20) ~ Ho*Range+E_Latitudeabs*Range+I(E_Latitudeabs^2) + (1|q), data= pcho_mean),  keep.effs="Ho", reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")
m<-lmer(log(T_seed_20)~ Ho+E_Latitudeabs+Range+I(E_Latitudeabs^2)  + (1|q) , data = pcho_mean)
visreg(m, by="Range",xvar="Ho", overlay=T)
interactionMeans(m, slope="Ho" )

m1<-lm(log(T_seed_20)~ Ho+E_Latitudeabs*Range , data = pcho_mean)
visreg(m1, by="Range",xvar="Ho", overlay=T)
interactionMeans(m1, slope="Ho" )

Anova(lmer(log(T_seed_20) ~ Ho+E_Latitudeabs+Range +I(E_Latitudeabs^2)+ (1|q) , data = pcho_mean),type =3, ddf="Kenward-Roger", test.statistic = "F")


step(lmer(log(T_sexmismatch+60) ~ Ho*Range+E_Latitudeabs*Range+I(E_Latitudeabs^2) + (1|q), data= pcho_mean),  keep.effs="Ho", reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

step(lmer(log(T_sexrat_weight+1) ~ Ho*Range+E_Latitudeabs*Range +I(E_Latitudeabs^2)+ (1|q), data= pcho_mean),  keep.effs="Ho", reduce.fixed = TRUE, reduce.random = FALSE, alpha.fixed = 0.05,type =3, ddf="Kenward-Roger")

tpm =read.table("tmp.txt")
tpm2= p.adjust(tpm$V1, method = "fdr")
write.table(tpm2, "tmp2.txt")
```


**END SCRIPT**
===